---
title: "MA615 Unit 1 Final Assignment"
author: "Rong Li"
date: "2020/10/19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("tidyverse","knitr","kableExtra","ggplot2")
options (warn = -1)
```

# 1 Summary
In this assignment, we use the data set berries from USDA. We focus on one kind of berries--strawberries. Finish cleaning and organizing the data. Visualize and explore the data. This report can be divided into 2 parts: Data cleaning and EDA. 

* Data cleaning: We select the data of strawberries from the raw data. Seperate some columns and eliminate the redundancy.
* EDA: We plot the boxplots for values of every year and of every states. Then we conduct the pca for values.

Besides, we deploy a shiny app [Shiny_strawberries](https://rongli.shinyapps.io/shiny/) for data display.

# 2 Data cleaning
In this part, we obtain the data set strawberries that we will conduct EDA for later.

## 2.1 Deal with the raw data set berries
First, we explore the data set berries. Find out the redundent columns and delete them. Then display the data set.  

```{r}
## read the data
ag_data <- read_csv("/Users/amelia/Documents/mssp/MA615/Hw_berries/Berries_strawberries/data/berries.csv", col_names = TRUE)

## look at number of unique values in each column
aa <- summarize_all(ag_data, n_distinct)

## make a list of the columns with only one unique value
bb <- which(aa[1,] == 1)

## list the 1-unique valu column names 
colnames(ag_data)[bb]

## remove the 1-unique columns from the dataset
ag_data <- select(ag_data, -all_of(bb))
aa <- select(aa, -all_of(bb)) 

## State name and the State ANSI code are (sort of) redundant
## Just keep the name
ag_data <- select(ag_data, -4)
aa <- select(aa, -4) 

## Display the head of ag_data 
head(ag_data)
```

## 2.2 Get the raw data of strawberries 
Now, we focus on one kind of berries--strawberries.

```{r}
## Strawberries
sberry <- filter(ag_data, (Commodity=="STRAWBERRIES") & (Period=="YEAR"))
sberry <- select(sberry, -c(Period, Commodity))
```

## 2.3 Separate the column 'Data Item' into 3 columns "type", "meas" and "what".

```{r}
#### Does every Data Item begin with "STRAWBERRIES, "
sum(str_detect(sberry$`Data Item`, "^STRAWBERRIES, ")) == length(sberry$`Data Item`)

## Seperate the Data Item
sberry1 <- subset(sberry, str_detect(sberry$`Data Item`, "^STRAWBERRIES, ") == "TRUE")
sberry2 <- subset(sberry, str_detect(sberry$`Data Item`, "^STRAWBERRIES, ") == "FALSE")
sberry1 <- separate(sberry1, `Data Item`, c("B","type", "meas", "what"), sep = ",") 
sberry2 <- separate(sberry2, `Data Item`, c("B","todo"), sep = "-") 
sberry2 <- separate(sberry2, `todo`, c("type","meas", "what"), sep = ",") 
sberry <- rbind(sberry1, sberry2)
sberry <- select(sberry, -B)
```

## 2.4 Seperate the column "type" into 3 columns "type", "lab1" and "lab2".

```{r}
## Seperate the type
sberry1 <- subset(sberry, str_detect(sberry$`type`, " - ") == "TRUE")
sberry1 <- separate(sberry1, type,c("type", "lab1", "lab2"), " - ")
sberry1 <- separate(sberry1, type,c("b1", "type"), " ")

sberry2 <- subset(sberry, type == " FRESH MARKET")
sberry2$type <- "FRESH MARKET"

sberry3 <- subset(sberry, type == " PROCESSING")
sberry3$type <- "PROCESSING"

sberry4 <- subset(sberry, str_detect(sberry$`type`, " - ") == "FALSE" & 
                    type != " FRESH MARKET" & type != " PROCESSING")
sberry4 <- separate(sberry4, type,c("b1", "lab1", "lab2"), " ")

sberry <- plyr::rbind.fill(sberry1, sberry2, sberry3, sberry4)

sberry[is.na(sberry)] <- ""

sberry <- select(sberry, -b1)
## OK now Data Item has been split into parts
```

## 2.5 Separate the column "Domain" into 2 columns "D_left" and "D_right".

```{r}
## onto Domain
sberry <- separate(sberry, Domain, c("D_left", "D_right"), sep = ", ")

sberry[is.na(sberry)] <- ""
```

## 2.6 Separate the column "Domain Category" into 4 columns "DC_left_l", "DC_left_r", "DC_right_l", "DC_right_r".

```{r}
## And now Domain Category
sberry <- separate(sberry, `Domain Category`, c("DC_left", "DC_right"), sep = ", ")

## unique(sberry$DC_left)
## unique(sberry$DC_right)

## looks like DC_left combines labels
sberry$DC_right[which(str_detect(sberry$DC_left,"138831"))]="INSECTICIDE: (CYFLUMETOFEN = 138831)"
sberry$DC_left[which(str_detect(sberry$DC_left,"138831"))]="CHEMICAL"


## work on DC_left first

sberry <- separate(sberry, DC_left, c("DC_left_l", "DC_left_r"), sep = ": ")

## sberry$DC_left_l %>% unique()
## sberry$DC_left_r %>% unique()

## now work on DC_right

sberry <- separate(sberry, DC_right, c("DC_right_l", "DC_right_r"), sep = ": ") 


sberry[is.na(sberry)] <- ""
```

## 2.7 Eliminate the redundancy in data set strawberries
a) remove column "DC_left_l"; b) remove column "DC_right_l"; c)remove

```{r}
## fine and remove redundant columns

## Test for D_left, DC_left_l
## paste(sberry$D_left,sberry$DC_left_l) %>% unique
## returns -- "CHEMICAL CHEMICAL"     "FERTILIZER FERTILIZER" "TOTAL NOT SPECIFIED" 

## remove column bberry$DC_left_l
sberry <- select(sberry, -DC_left_l)

## Test for D_right, DC_right_l
# sum(sberry$D_right == sberry$DC_right_l)
# [1] 3220
# sberry$DC_left_r %>% unique()
# [1] ""            "(NITROGEN)"  "(PHOSPHATE)" "(POTASH)"    "(SULFUR)"   

## remove column DC_right_l
sberry %<>% select(-DC_right_l)
```























